---
title: "DAVID Redone Code"
author: "Anusuiya Bora; s222331482@deakin.edu.au"
#output: html_document
output: 
  html_document:
    code_folding: show 
date: "2024-04-16"
---

### Methodology:
1. Loading libraries

```{r loading_libraries}

#install.packages("pacman")  #this installs package. Only do once.

library("clusterProfiler")
library("kableExtra")
library("dplyr")
library("rio")
library("here")
library("gtsummary")
library("ggplot2")


# pacman::p_load(clusterProfiler,
#               kableExtra,
#               rio,           # package for importing and exporting data
#               here,          # points R in particular directories to look for files; uses relative file paths; to tell R where our data is; used with "rio" package; good for relative file path
#               dplyr,         # to manipulate data
#               usethis,       # handy data cleaning functions
#               gtsummary,
#               plotly,
#               ggplot2
#)

```


2) Input gene list:

``` {r loading_input_genes}

genelist <- read.csv("PMC6587650.csv")
Genevec <- genelist[, 1]

```

3) Loading GO datsets

```{r include=FALSE, echo=FALSE, load_GORDS}

gobp <- rio::import(here::here("DAVID_Reborn", "GO_dataset", "GObp.rds"))
gocc <- rio::import(here::here("DAVID_Reborn", "GO_dataset", "GOcc.rds"))
gomf <- rio::import(here::here("DAVID_Reborn", "GO_dataset", "GOmf.rds"))
gocombined <- rio::import(here::here("DAVID_Reborn", "GO_dataset", "GOcombined.rds"))

#gobp <- readRDS("GObp.rds")
#gocc <- readRDS("GOcc.rds")
#gomf <- readRDS("GOmf.rds")
#gocombined <- readRDS("GOcombined.rds")

```


```{r include=FALSE, echo=FALSE, converting_RDS2gmt}
gobp_terms <- rev(stack(gobp))
str(gobp_terms)

gocc_terms <- rev(stack(gocc))
str(gocc_terms)

gomf_terms <- rev(stack(gomf))
str(gomf_terms)

gocombined_terms <- rev(stack(gocombined))
str(gocombined_terms)

#goterms <- rev(stack(golist))
#str(goterms)

```

5) Performing Enrichment analysis

```{r include=T, echo=T, enricher_code}
ora <- as.data.frame(enricher(gene = Genevec ,
                              minGSSize = 10,
                              maxGSSize = 50000, 
                              TERM2GENE = gocombined_terms, #for TERM2GENE, options are: gobp_terms, gocc_terms, gomf_terms, gocombined_terms
                              pAdjustMethod="fdr",  
                              pvalueCutoff = 1, 
                              qvalueCutoff = 1  ))

x <- gsub("/", " ", ora$geneID) #to separate geneIds from /

head(x)

dim(ora)

ora$geneID <- x

ora$ID <- NULL
rownames(ora) <- NULL #removing empty 1st column name

ora$S_No <- seq_len(nrow(ora))
ora <- ora[, c("S_No", names(ora)[-which(names(ora) == "S_No")])]

head(ora)

```


6) Calculating enrichment scores

``` {r include=FALSE, echo=FALSE, ES_calculation}

ora$gene <- NULL

gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

es <- gr/br

ora_new <- cbind(ora, gr, br, es)

names(ora_new)[names(ora_new) == "es"] <- "EnrichmentScore"

ora_arranged <- ora_new %>% 
  arrange(p.adjust)

head(ora_arranged)


```

7) Selecting pathways with FDR<0.05

```{r pathways_FDR}

ora_fdr <- ora_arranged %>%
  filter(p.adjust < 0.05)

```



```{r output_table}

ora_fdr %>%
  kbl(caption="Genes and GO Terms identified") %>%
  kable_paper("hover", full_width = F)

```

```{r complete_results, results='asis', echo=FALSE}

HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )
 
cat(HEADER)
 
kbl(ora_arranged, format="html", row.names=FALSE, caption = "Complete results",digits=100) %>%
kable_styling("hover", full_width = FALSE)
 
cat("<br></p></details>")
cat("<hr><br>")

```

```{r pie_chart}

aggregate_data <- aggregate(EnrichmentScore ~ Description, data = ora_fdr, FUN = sum)

# Create a pie chart
pie(aggregate_data$EnrichmentScore, labels = aggregate_data$Description)

ggplot(ora_fdr, aes(x = "", y = Description, fill = EnrichmentScore)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Pie Chart for enriched terms", fill = "EnrichmentScore", y = NULL) +
  theme_minimal()

```

### Save object

```{r save_object}

saveRDS(ora_arranged, file = "ora.RDS")
# z <- readRDS("ora.RDS")  #To read RDS 

```

### Session information

For reproducibility.

```{r,sessioninfo}

sessionInfo()

```
